---
title: "Standard ASE Analysis"
author: "Lynn Ly"
date: "3/23/2018"
output: html_document
---

**Part 1: Parental Bias** Takes the genes determined have differential allelic imbalance between F1_414 and F1_415. Classifies them according to which parent they are biased toward.  
**Part 2: Regulatory Effect** Looks at all genes with exonic SNVs. Classifies based on being regulated by cis, trans, both, or neither. Counts were normalized with DESeq2 size factors.  
*cis* test: binomial exact test on F1_414_Ae and F1_414_Ol counts, centered on the global reference bias.  
*trans* test: fisher's exact test on (parent) Ae and Ol counts compared to (F1) Ae and Ol counts.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 
```{r}
load("phasedData.Rdata")
load("Results_Beta/significantb.Rdata")

MBASED_results <- significantb$geneOutput %>% 
                  mutate(GeneID = rownames(significantb$geneOutput)) %>%
                  select(GeneID, MAFD = majorAlleleFrequencyDifference, p = pValueASE)

parentOfOrigin <- phasedData %>% 
                  dplyr::select(GeneID, F1_414_RO, F1_414_AO, F1_415_RO, F1_415_AO) %>%
                  group_by(GeneID) %>% 
                  summarise_at(vars(F1_414_RO:F1_415_AO), sum)

parentOfOrigin %<>% filter(GeneID %in% MBASED_results$GeneID) %>%
                    left_join(MBASED_results) %>%
                    mutate(F1_414_mat = signif(F1_414_RO / (F1_414_RO + F1_414_AO), 3),
                           F1_414_pat = signif(F1_414_AO / (F1_414_RO + F1_414_AO), 3), 
                           F1_415_mat = signif(F1_415_AO / (F1_415_RO + F1_415_AO), 3), 
                           F1_415_pat = signif(F1_415_RO / (F1_415_RO + F1_415_AO), 3))

parentOfOrigin %<>% mutate(mat = F1_414_mat + F1_415_mat,
                          pat = F1_414_pat + F1_415_pat) %>% 
                    mutate(Parent = ifelse(mat > pat, "Maternal", "Paternal"))
head(parentOfOrigin)

imprintedGenesF1 <- parentOfOrigin
save(imprintedGenesF1, file = "imprintedGenesF1.Rdata")
```

### Cis vs trans regulatory effect
Using only F1_414
```{r}
bt <- function(a, b, p = 0.5) {binom.test(a, a + b, p)$p.value}
ft <- function(a, b, c, d) {
  df <- data.frame(Parent = c(a, b), F1 = c(c, d))
  fisher.test(df)$p.value
}

AddEffectSize <- function(regulatoryData) {
  regulatoryData %<>%
      mutate(parentAI = log2(Ae / Ol), 
             F1AI = log2(F1_Ae / F1_Ol), 
             diff = parentAI - F1AI)
  return(regulatoryData)
}

NormalizeBySizeFactor <- function(regulatoryData) {
  sizeFactors <- DESeq2::estimateSizeFactorsForMatrix(regulatory[2:5])
  regulatoryData %<>%
    mutate(Ae = Ae / sizeFactors['Ae'], Ol = Ol / sizeFactors['Ol'],
           F1_Ae = F1_Ae / sizeFactors['F1_Ae'], F1_Ol = F1_Ol / sizeFactors['F1_Ol']) %>%
    mutate_at(vars(2:5), round)
  return(regulatoryData)
}

ClassifyRegulatoryEffect <- function(regulatory, p = 0.05) {
  regulatory$effect <- mapply(ClassifyRegulatoryEffectHelper, regulatory$pCis, regulatory$pTrans)
  return(regulatory)
}

ClassifyRegulatoryEffectHelper <- function(pCis, pTrans, p = 0.05) {
  if(pCis < p & pTrans < p) {
    return("cis and trans")
  } else if (pCis < p) {
    return("cis")
  } else if (pTrans < p) {
    return("trans")
  } else {
    return("conserved")
  }
}

FindRegulatoryEffect <- function(phasedData, groupByGene = TRUE, effectSize = FALSE) {
  regulatory <- phasedData %>% select(GeneID, Ae_RO, Ae_AO, Ol_RO, Ol_AO, F1_414_RO, F1_414_AO, F1_414_refBias)
  if(groupByGene == TRUE) {
    sums <- regulatory %>% group_by(GeneID) %>%
      summarise_at(vars(Ae_RO:F1_414_AO), sum)
    refBias <- regulatory %>% group_by(GeneID) %>%
      summarise(F1_414_refBias = mean(F1_414_refBias))
    regulatory <- full_join(sums, refBias)
  }

  regulatory %<>% transform(Ae = Ae_RO + Ae_AO, Ol = Ol_RO + Ol_AO) %>%
                  select(GeneID, Ae, Ol, F1_Ae = F1_414_RO, F1_Ol = F1_414_AO, refBias = F1_414_refBias) 
  
  regulatory %<>% NormalizeBySizeFactor()
  
  if(effectSize == TRUE) {
    regulatory <- AddEffectSize(regulatory)
  }
  
  regulatory$pCis <- mapply(bt, regulatory$F1_Ae, regulatory$F1_Ol, regulatory$refBias)
  regulatory$pTrans <- mapply(ft, regulatory$Ae, regulatory$Ol, regulatory$F1_Ae, regulatory$F1_Ol)
  
  regulatory$pCis <- p.adjust(regulatory$pCis, method = "bonferroni")
  regulatory$pTrans <- p.adjust(regulatory$pTrans, method = "bonferroni")
  
  return(regulatory)
}

genewise <- FindRegulatoryEffect(phasedData)

regulatoryEffectsF1 <- genewise %>% select(-refBias) %>%
  ClassifyRegulatoryEffect(p = 0.05)

save(regulatoryEffectsF1, file = "regulatoryEffectsF1.Rdata")
```
