---
title: "aseQTL subset"
author: "Lynn Ly"
date: "3/21/2018"
output: html_document
---

**I don't think this workflow will work for ASE and an intercross.**


I have 18222 loci with informative SNVs for potential aseQTL analysis. Looking at that will be computationally expensive, so this is for preliminary aseQTL on a random subset, using only 100 SNVs.  

Quality control section is excluded, since that was performed on the whole dataset. 

```{r setup, include=FALSE}
library(tidyverse)
library(qtl)
library(snow)
```

### Prep
Subset 100 SNVs
```{r}
aseRatiosF2_wide <- read.csv(file = "../input/aseRatiosF2_wide.csv")

set.seed(1024)
aseRatiosF2_subset <- sample_n(aseRatiosF2_wide, size = 100)

write.csv(aseRatiosF2_subset, file = "../input/aseRatiosF2_subset.csv", row.names = FALSE)
```

Read in data from the marker file (genetic map) and my phenotypes (the ASE ratios) 
```{r}
data.qtl <- read.cross(format = "csvsr", 
                       genfile = "../input/LG.f2.madmapper.final_gen_revised_flipped.csv", 
                       phefile = "../input/aseRatiosF2_subset.csv",
                       genotypes = c("AA", "AB", "BB")
                       )
```

### QTL on Subset

Create imputations and calculate the conditional genotype probabilities
```{r}
data.qtl <- sim.geno(data.qtl, step = 1, n.draws = 32) # imputation?  
data.qtl <- calc.genoprob(data.qtl, step=1) 
```

(example) scan one via multiple imputation method
```{r}
system.time(out.imp <- scanone(data.qtl, pheno.col = 2:ncol(data.qtl$pheno),
                   method = "imp", use = "all.obs"))
```

### Significance Thresholds
How large of a LOD score do you need to call it as significant? Null hypothesis is derived from permutation tests: permute the phenotypes relative to the genotype data. Then, the QTL mapping method is applied to the shuffled version to obtain a set of LOD curves. The genome-wide maximum LOD score is derived from these curves. 

```{r}
operm <- scanone(data.qtl, pheno.col = 2:ncol(data.qtl$pheno), method = "imp",
                 use = "all.obs", n.perm = 1000, n.cluster = 8)

summary(operms, alpha = c(0.20, 0.05)) # gives the LOD thresholds for 80% and 95% confidence intervals

summary(out.imp, perms = operms, alpha = 0.05) # picks out the LOD peaks that meet 5% significance level

save(data.qtl, out.imp, operms, file = "results/scanone_aseQTL_subset.Rdata")
```

```{r}
load("results/scanone_aseQTL_subset.Rdata")
```

